\documentclass[sigconf]{acmart}

\usepackage{xspace,listings,url,framed,amssymb,colortbl,
       amsmath,mathpartir,hyperref,doi, rotating, stmaryrd, 
       graphicx, tikz, colortbl, xparse, etoolbox, pgffor,booktabs} 
\input{macros}
\newcommand{\code}[1]{{\tt #1}\xspace}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\lstdefinelanguage{JavaScript}{
  keywords={typeof,new,true,false,instanceof,catch,function,return,null, 
    catch, switch, var, if, in, while, do, else, case, break},
  keywordstyle=\color{darkgray},
  ndkeywords={class,def,interface,export,boolean,throw,extends,implements,import,this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,  comment=[l]{//},  morecomment=[s]{/*}{*/},
  commentstyle=\color{gray}\ttfamily,  stringstyle=\color{gray}\ttfamily,
  morestring=[b]',  morestring=[b]",
  backgroundcolor=\color{vlightgray},
  aboveskip=\medskipamount, %0em,
  belowskip=\smallskipamount, %0em
  escapeinside={(*@}{@*)}
}
\lstset{
  language=JavaScript,  extendedchars=true,  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,   showspaces=false,  numberstyle=\small,
  numbersep=9pt,  tabsize=2, breaklines=true,  showtabs=false, captionpos=b
}

% DOI
%\acmDOI{10.475/123_4}

% ISBN
%\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[FtFJP]{Submitted to FtFJP}{July 2018}{}
\acmYear{2018}
%\copyrightyear{2016}
%\acmArticle{4}
%\acmPrice{15.00}

\begin{document}
\title{Monotonic Gradual Typing in a Common Calculus}
\subtitle{}

\author{Benjamin Chung}
\affiliation{\institution{Northeastern University}}
\author{Jan Vitek}
\affiliation{\institution{Northeastern and CTU}}


\begin{abstract}
Gradual typing refers to the notion that programs can be incrementally
decorated with type annotations. Languages that
support this approach to software development allow for programs being in
various state of ``typedness'' on a scale ranging from entirely untyped
to fully statically typed. The intellectual challenge for language designers
is to assign meaning programs on that scale. The litterature in the field is
rich with a variety of proposals ranging from the laissez faire of optional
typing to sound approaches that wrap all values in protective shells to
ensure that they behave as promised by their ascribed type. Each gradual
type system comes with tradeoffs. Some systems provide strong guarantees at
the expense of vastly degraded performance, others do not impact the running
time of programs but do little to pervent type errors. This paper looks at
an intriguing point the landscape: a variant of Python referred to as
Monotonic Reticulated Python. Our contribution is a semantics of the
proposal as a translation from a surface language down to a core calculus.
\end{abstract}

\definecolor{Gray}{gray}{0.9}

\definecolor{vlightgray}{gray}{0.93}

\maketitle

\section{Introduction}

Gradual typing refers to a family of approaches that add types to untyped
programs. These approaches each protect the boundary between typed and untype
code in different ways~\cite{SiekTaha06,tf-dls06}. Gradual typing allows
untyped code to pass values into typed code, and the static types need to be
reconciled with the values dynamically. The pantheon of gradual type systems
arises from different mechanisms for checking the types of untyped values.

Object-oriented languages pose a particular problem for gradual typing. It is
easy to check if a base value has a type, but problems arise from higher-order
behavior (that is, programs that pass around behavior inside values). In a
functional language, this manifests when trying to assert an arrow type on a
lambda, whose return type cannot be checked at the cast site. Object-oriented
languages have this same problem -- trying to assert a typed method on an
object with an untyped version of the method -- but have additional complexity
in the form of subtyping.

\begin{figure}[h]
\begin{lstlisting}
class C { 
  f(x:*):* { x } 
  g(x:*):* { new C() }
}
class D { 
  f(x:D):D { x }
}

new D().f( <*> new C() )
\end{lstlisting}
\caption{A gradually-typed program.}
\label{fig:simpleprog}
\end{figure}

Figure~\ref{fig:simpleprog} shows a simple example of a gradually typed
object-oriented program. Typed expressions use a standard Featherweight Java-
like type system with structural subtyping but not all expressions in this
program are typed. In the surface language, the type \any indicates that a
variable is dynamic; thus it can hold a value of any class, and any method can
be invoked with that variable as the receiver, returning a value of type \any.
The program uses dynamic behavior (a cast) to pass an instance of class \C as
one of type \D in the call to \f. \C has no guarantee to act like a \D, so
\D-ness must be enforced at runtime. However, the best way to enforce static
types on untyped values is an open problem. The program shown here could fail
when the instance of \C is first passed, fail if \f is called on \x, or not
fail at all if it does not violate the guarantee enforced by the gradual type
system.

In a companion paper, to appear at ECOOP 2018~\cite{us}, we examined four
gradual typing approaches, each of which addresses this problem differently.
In that paper, we discussed the optional, transient, behavioral, and concrete
semantics. We chose these semantics due to their popularity and relative
simplicity; but elided the monotonic semantics of~\cite{siek14}, due to its
complexity. This paper places the monotonic semantics into the same framework
as the companion paper.

The monotonic semantics sets out to solve a key problem for gradual typing:
speed of field access. For example, the behavioral semantics adds a wrapper
around every field access which checks the values type, at the cost of
indirection. This can make typed code in a gradually typed language
excruciatingly slow~\cite{popl16}.  Suppose we have an object $\x = \{ \f:\any
= 3 \}$, then alias it as $\xp = \SubCast{\{ \f:\xt{int}   \}}\x$. Despite
\xp's static type, the field dereference $\xp.\f$ could return a value of any
type. If we run \xt{\x.\f = "hello"} at some point, then $\xp.\f$ is now ill-
typed via the reference. As mentioned previously, other semantics deal with
this by checking the result of operations in typed code, lazily enforcing
soundness of type annotations.

The monotonic semantics is an \emph{eager} mechanism to enforce types,
checking values assigned to fields on assignment rather than dereference. With
monotonic, once a value in the heap is typed, it will always remain of that
type. In the example above, once we aliased \x as \xp, \xp's type (e.g. that
\f has type \xt{int}) will be enforced upon usages of \x. Now, our assignment
\xt{\x.\f = "hello"} will fail, because we are assigning a value of improper
type to the field \f of the object. Monotonic uses this mechanism to guarantee
that all values referred to by typed references will be of that type.

In this paper, we present a formalization of the monotonic semantics in the
same framework used in our companion paper~\cite{us} for comparing gradual
type systems. We use a common gradually-typed object-oriented source language
and translate it to a common target language (called \kafka) according to the
monotonic semantics. We then use this formalization to discuss the monotonic
semantic properties, and compare it to the gradual type systems presented
in our companion paper.

\section{Monotonic Gradual Types}

The idea behind the monotonic semantics was originally put forward by Siek
and Vitousek in~\cite{siek14} for their Reticulated Python language
implementation. The same group then went on to implement the monotonic
semantics into their Grift implementation (a functional language), allowing
the comparison of performance inherent to each gradual type system.

The idea behind the monotonic semantics is that types only ever get more
precise. This design gives rise to its name; types of values in a monotonic
language's heap monotonically get more static. Each object in the heap has
an associated type, it's run-time tag, in an object-oriented language that
tag identifies the class of the object and is used for dispatching method
invocations. When a reference to an object is cast to a more precise type,
intuitively a type with fewer occurrences of \any, that new type is attached
to the value in the heap.  If untyped code later tries to read or write
fields, or call methods, in a way that violates the updated type, a type
error will occur. By locking in applied types forevermore, typed code under
the monotonic semantics can assume that mutable state will remain well
typed.

In turn, the ability to rely on the types of mutable state 
provides several advantages over other gradual type systems. The monotonic
system has a soundness guarantee, unlike the optional semantics (which
simply forgets about all type annotations at runtime); monotonic languages
can elide field dereference guards needed by the transient semantics; do not
need the wrappers inserted by the behavioral semantics; and is far more
flexible than the concrete semantics. Additional details about these other
semantics can be found in our companion paper~\cite{us}.

As a result, monotonic has the potential to provide substantial benefits to
the performance of gradually typed languages. Variables are guaranteed to
hold values of their cast type, so they can be relied upon for memory layout
of objects. Types under the monotonic semantics can be used to optimize away
generic field accesses, and assignments to typed fields from fully typed
code do not need to be checked.

In the context of object orientation, monotonic's ability to protect mutable
state without the need for wrappers is a significant advantage. Other
gradual type systems need guards, wrappers, or a stronger notion of
soundness in order to guarantee safety of field access. The monotonic
semantics, in contrast, prevents untyped code from assigning ill-typed
values to previously typed references, allowing unchecked dereferencing in
typed code.

\section{Formalism}

We use the framework we set out in our companion paper to provide a simple
presentation of the monotonic semantics in an object-oriented setting. Instead
of using formal languages specially crafted to support each semantics, we use
a common source and target language for every gradual typing semantics. The 
semantics is then encoded into the translation, allowing comparison of gradual 
type systems by examining their translation.

\subsection{Source Language}

Our common source language is similar to Featherweight Java, with a few
changes. This language has classes but relies on structural subtyping
instead of the nominal subtyping adopted by FJ. For our purposes there is no
need for inheritance. Structural subtyping is needed to support the
evolution of types as casts are being applied. The source language also
introduces a dynamic type, \any, and a convertibility relation, allowing
gradually typed programs to be expressed. This same source language and
source language type system is shared between the translation presented here
and the four in our main paper, allowing the same programs to be ran under
all of our semantics. We replicate the static semantics for the source
language here, in figure~\ref{slts}. The dynamic of the source language is
given by a translation to a lower-level language which we will present next.

Programs in the source language consist of a class table, \K, and an initial
expression, \e. Classes consist of field and method definitions. Fields have
a name and a type, \f:\t. Method definitions have names, types for their
single argument and return type, and an expression. Expressions consist of
variables, the self-reference \this, field read and write \FRead\f and
\FWrite\f\e, invocation \Call\e\m\e, and new object creation
\New\C{\e[1]..}. Each class has an implicit default constructor that merely
assigns arguments to fields. All operations conform to what they would do in
Featherweight Java, and the type system is straightforward.

The static type system allows the implicit conversion of typed and untyped
terms via the convertibility operator, written $\ConvertE\K s\t\tp$ -- type
$\t$ is convertible to type $\tp$.  The relation is used both for up-casting
and for conversions of \any to non-\any types.  \RuleRef{SUB} ensures
soundness of up-casts. While \RuleRef{TOA} and \RuleRef{ANYC} allow implicit
conversion to and from the dynamic type.  To avoid collapsing the type
hierarchy, convertibility is not transitive. \RuleRef{TOA} and
\RuleRef{ANYC} are break soundness, their use motivates the need for dynamic
checks. 

The source language enforce privacy on fields. They can only be accessed
from them method of the object. Furthermore, fields do not show up in type
signatures. Subtyping between objects is computed solely on methods
signatures.


\begin{figure*}[!t]

\hrulefill

\vspace{4mm}

\begin{tabular}{@{}l@{~}l@{}l@{}l@{}l@{}l@{}l@{}l}
\e~::=~\x\B\this\B\FRead\f\B\FWrite\f\e\B\Call\e\m\e\B\New\C{\e[1]..}\\
\k~::=~ \Class \C {\fd[1]..}{\md[1]..} \qquad
\md~::=~\Mdef\m\x\t\t\e\qquad
\fd~::=~ \Fdef\f\t\qquad
\t~::=~ \any \B \C
\end{tabular}

\begin{mathpar}
\Rule{STG-VAR}{~\\\\
  \HasType \Env\x\t
}{
  \EnvTypeS \Env\K\x\t
}

\Rule{STG-GET}{
  \HasType \Env\this\C \\\\  \Fdef\f\t \in \App\K\C
}{
  \EnvTypeS \Env\K{\FRead\f}\t
}    

\Rule{STG-SET}{
  \HasType \Env\this\C \quad \Fdef\f\t \in \App\K\C \\\\
  \EnvTypeS \Env\K\e\tp \quad  \ConvertE\K{s}\tp\t
}{
  \EnvTypeS \Env\K{\FWrite\f\e}\t
}    

\Rule[width=15em]{STG-CALL}{
  \EnvTypeS \Env\K\e\any \\\\ \EnvTypeS \Env\K\ep\t 
}{
  \EnvTypeS \Env\K{\Call\e\m\ep}{\any}
}    

\Rule[width=15em]{STG-CALL}{
  \EnvTypeS \Env\K\e\C \quad \EnvTypeS \Env\K\ep\t \\\\
  \Mtype \m{\t[1]}{\t[2]}\in \App\K\C  \quad
  \ConvertE\K{s}\t{\t[1]}
}{
  \EnvTypeS \Env\K{\Call\e\m\ep}{\t[2]}
}   

\Rule{STG-NEW}{
  \Ftype{\f[1]}{\t[1]}.. \in \App\K\C \\\\
  \EnvTypeS \Env\K{\e[1]}{\tp[1]}..\quad \ConvertE\K{s}{\tp[1]}{\t[1]}..
}{
  \EnvTypeS \Env\K{\New\C{\e[1]..}}\C
}
\end{mathpar}

\vspace{-5mm}  
  
\begin{mathpar}
\IRule{SUB}{
   \SSub\emptyset\K\t\tp
}{
   \ConvertE\K{s}\t\tp
}
    
\IRule{TOA}{~ }{ \ConvertE\K{s}\t\any}
    
\IRule{ANYC}{~}{ \ConvertE\K{s}\any\t }
\end{mathpar}


\hrulefill
\caption{Surface language syntax and type system (extract).}\label{slts}
\end{figure*}

\begin{figure*}[!t]

\hrulefill

\vspace{4mm}

\small\begin{tabular}{@{}lll}

\begin{minipage}{5.5cm}\begin{tabular}{@{}l@{~}l@{}l@{}l@{}l@{}l@{}l@{}l}
\e\hspace{.1cm} ::= & \hspace{.2cm} \x        
   &\B \this         
   &\B \that      
   &\B \FRead\f  \\   
   & &
   &\B \FWrite\f\e 
   &\B \KCall\e\m\e\t\t \\
   & &   
   &\B \New\C{\e[1]..}  
   &\B \DynCall\e\m\e \\
   & & & \multicolumn{3}{@{}l}{\B \a \B \FReadR\a\f \B \FWriteR\a\f\e} \\
   & & & \B \src{\MonCast\t\e}
\end{tabular}\end{minipage}&
\begin{minipage}{3.7cm}\begin{tabular}{l@{~}l@{}l@{}l}
  \k &::= \Class \C {\fd[1]..}{\md[1]..} \\
 \md &::= ~ \Mdef\m\x\t\t\e \\ 
 \fd &::= ~ \Fdef\f\t \\ 
  \t &::= ~ \any  \B   \C  \\ 
\end{tabular}\end{minipage} &
\begin{minipage}{5.5cm}

\begin{tabular}{llllllllllllllllll}
\EE &::=& ~ \FWriteR\a\f\EE   &\B  
        \KCall\EE\m\e\t\t  &\B\\
        &&\KCall\a\m{\EE}\t\t &\B
        \DynCall\EE\m\e   &\B\\
&       & \DynCall\a\m\EE   &\B
       \src{\MonCast\t\EE}  &\B\\
&&     \New\C{\a[1]..\,\EE\,\e[1]..} &\B 
      \EM{\square}
\end{tabular}
\end{minipage}
\end{tabular}


\begin{mathpar}
\IRule{KT-VAR}{
   ~\\\\
   ~\\\\
   \HasType \Env\x\t
 }{
   \EnvType \Env\s\K\x\t
}

\IRule{KT-SUB}{
  ~\\\\
  \EnvType \Env\s\K\e\tp \\\\
 \StrSub \cdot\K \tp \t
 }{
  \EnvType \Env\s\K\e\t 
}   

\IRule{KT-READ}{
  ~\\\\
  \HasType\Env\this\C \\\\
  \Fdef\f\t \in \App\K\C
}{
  \EnvType \Env\s\K{\FRead\f}\t
}  

\IRule{KT-REFREAD}{
  \s(\a) = \C\{..\} \\\\
  \Fdef\f\t \in \App\K\C
}{
  \EnvType \Env\s\K{\FReadR\a\f}\t
}  

\IRule{KT-WRITE}{
  \HasType\Env\this\C \\\\
  \Fdef\f\t \in \App\K\C \\\\
  \EnvType \Env\s\K\e\t
}{
  \EnvType \Env\s\K{\FWrite\f\e}\t
}    

\IRule[width=12em]{KT-REFWRITE}{
  \s(\a) = \C\{..\} \\\\
  \Fdef\f\t \in \App\K\C \\\\
  \EnvType \Env\s\K\e\t
}{
  \EnvType \Env\s\K{\FWriteR\a\f\e}\t
}  

\IRule[width=16em]{KT-CALL}{
  \EnvType \Env\s\K\e\C \\\\
  \EnvType \Env\s\K\ep\t \\\
  \Mtype\m\t\tp \in \App\K\C 
}{
  \EnvType \Env\s\K{\KCall\e\m\ep\t\tp}\tp
}    

\IRule{KT-DYNCALL}{
  ~\\\\
  \EnvType \Env\s\K\e\any \\\\
  \EnvType \Env\s\K\ep\any
}{
  \EnvType \Env\s\K{\DynCall\e\m\ep}\any
}    

\IRule[width=20em]{KT-NEW}{
  ~\\\\
  \EnvType \Env\s\K{\e[1]}{\t[1]}..\\\\
  \Class \C {\Fdef{\f[1]}{\t[1]}..}{\md[1]..} \in \K
}{
  \EnvType\Env\s\K{\New\C{\e[1]..}}\C
}

\src{
\IRule{KT-MONOCAST}{
  \EnvType \Env\s\K\e\tp
}{
  \EnvType \Env\s\K{\MonCast\t\e}\t
}
}

\IRule{KT-REFTYPE}{
  \s(\a) = \obj\C{\ap[1]..}
}{
  \EnvType \Env\s\K\a\C
}

\IRule{KT-REFANY}{
 }{
   \EnvType \Env\s\K\a\any
}
\end{mathpar}

\vspace{4mm}

\begin{minipage}{0.45\textwidth}
\begin{tabbing}
\K\HS\New\C{\a[1]..}\HS\=\s~\HS\=\Red\HS\=\K\HS\=\ap\HS\=\sp\HS\=
   \WHERE\HS\=\fresh\ap \HS\HS\sp ={\Map\s{\Bind\ap{\obj\C{\a[1]..}}}}
\\
\K\HS \FReadR\a{\f[i]} \> \s \>\Red\>  \K\>$\a[i]$ \> \s  \> 
   \WHERE \>\App\s\a=\obj\C{\a[1],..\a[i],\a[n]..}
\\
\K\HS{\FWriteR\a{\f[i]}\ap}\>\s\>\Red\>\K\>\ap\>\sp\>  
   \WHERE \>\App\s\a=\obj\C{\a[1],..\a[i],\a[n]..} 
   \HS\HS \sp=\Map\s{\Bind\a{\obj\C{\a[1],..\ap,\a[n]..}}}
\\
\K\HS{\KCall\a\m\ap\t\tp} \> \s      \>\Red\>     \K \>  \ep \> \s \> 
    \WHERE\> \ep = {[\a/\this~{\ap/\x}]\e} \HS\HS
             \Mdef\m\x{\t_{1}}{\t_{2}}\e\In \App\K\C \HS\HS
             \App\s\a=\obj\C{\a[1]..} \HS\HS
               \StrSub {\emptyset}\K\t {\t_{1}} \HS\HS
 \StrSub {\emptyset}\K{\t_{2}} \tp
\\
 \K\HS {\DynCall\a\m\ap}\> \s        \>\Red\>    \K \> \ep \> \s \>  
       \WHERE\> \ep = {[\a/\this~{\ap/\x}]\e}\HS\HS
                \Mdef\m\x\any\any\e \In \App\K\C \HS\HS
                 \App\s\a=\obj\C{\a[1]..} 
\\
 \K\HS {\MonCast \t\a} \> \s         \>\Red\>   \Kp \> \ap \> \sp \> \WHERE\> \moncast \a\t\s\K \Kp\ap\sp    
\\
\K \HS \EM{\EE[\e]} \> \s            \>\Red\>   \Kp \> \EM{\EE[\ep]} \> \sp \> \WHERE \> \K~\e~\s \Red~\Kp~\ep~\sp
\end{tabbing}\end{minipage}\vspace{2mm}

\noindent\hrulefill

\caption{\kafka syntax, static and dynamic semantics, and monotonic
  translation.}\label{syn}
\end{figure*}


\subsection{KafKa}

\kafka's design follows the surface language with some differences.
Expressions are extended with \that a distinguished variable used for
wrapped object reference, typed and untyped method invocations, subtype and
monotonic casts.  Evaluation is likewise standard with a few exceptions,
with an evaluation context consisting of a class table \K, an expression
being evaluated \e, and a heap \s, mapping from addresses \a to objects,
denoted $\C\{\a\ldots\}$. Due to the need for dynamically generate classes,
the class table \K is part of the state, and not the evaluation context.
\kafka has two invocation forms, the untyped invocation $\DynCall\e\m\ep$
and the typed invocation $\KCall\e\m\ep{\C}{\D}$, both denoting a method
call to method $\m$ with argument \ep. This is unusual for two
reasons. First, most gradual type systems have no explicit untyped
invocation form, and second, the typed form also specifies the argument type
and return type.
We use a gray background to indicate changes to \kafka in figure~\ref{slts} to
support monotonic. The only addition is that of the monotonic cast. The next
section focuses on giving a semantics to that cast.


\section{Monotonic Semantics}

The monotonic semantics is defined by a translation from the surface
language down to \kafka and the meaning of the monotonic cast. 

The translation is syntax directed, each surface class is translated to an
eponymous \kafka class. Types of methods and fields are retained in the
translation. Expressions are translated with $\TRG[M]{\e}\Env$. Variables
are translated to themselves. Calls are translated to either dynamic calls
or static calls depending on the type of the receiver.  If the receiver is
of type \any, then the argument must be of type \any. If the receiver of
some class \C then the argument must be of type \D, where \D is the expected
argument type of the method.  We use a variant of the translation relation
that performs a cast if necessary on its argument expression,
\TAG[M]{\e}{\Env}\t.  The inserted cast is the monotonic cast,
\MonCast{\t}{\ep}.

The semantics of monotonic arises from the implementation of the monocast
metafunction, used in the evaluation of monotonic casts shown in
figure~\ref{moncast}. At a high level, the cast operation computes the
meet of the source types (the most specific join of them), then applies it
by replacing the class associated with the object in memory. This is depicted
in rule \xt{CM}.

The first step in this operation is the meet operator. The meet operation is
akin to the one over a lattice, taking two types and always producing a type
higher on the type lattice. Two types have a meet if they are structurally
the same up to \any types distributed throughout their structure. The meet
itself consists of replacing all \any types in both types with the most
specific known type in that position from either type.

\begin{figure}[h]
\begin{lstlisting}
class C { m(x:*):* { x } n(x:C):C { x } }
class D { m(x:D):D { x } n(x:*):* { x } }
class E { m(x:D):D { x } n(x:C):C { x } }
\end{lstlisting}
\caption{Meet C and D}
\label{fig:meetex}
\end{figure}

\noindent
For example, if we were to compute the meet between \C and \D as  shown in
figure~\ref{fig:meetex}, we would end up with the type \xt{E}. The \any typed
argument to \m in \C is taken to be of type \D from class \D, while the
converse is used for \n in \D. The class \xt{E} then is more specific than
both \C and \D.

This operation relies on the meet metafunction which takes the two target
types, a context (for use in recursive types), and the class table. It then
outputs a new class, the meet of the two types, and an updated classes table
(meet can create multiple types). The context \xt{P}, either empty $\cdot$
or a context extended with a mapping $\P ~{{(\C,\D)}\mapsto\Cp}$, is used to
ensure termination of the meet operation on recursive types by storing the
names of meet-generated classes. Meet is deterministic, so meet of
previously-encountered types will be identical.  The meet operation can also
takes method definitions. It will pick out the comparable method definition
in the target method, and generate a suitable method. When two list of
methods are met, their result is passed into ifcgen, which generates an
``interface''. An interface is trivial class that exists only for the
purpose of giving us a type signature.  It will never ne

 with all of the method definitions obtained from meet. However, the
result of ifcgen is not a semantics-preserving class. 

Suppose we start out with $\x=\New\C{}$, then alias it with
$\xp=\MonCast\D\x$. If we then call $\DynCall\x\m{\New\D{}}$, and the
underlying object is an instance of \xt{E}, the dynamic call will fail as no
dynamically-typed method exists. As a result of this, our system does not
deem \xt{E} to be in a subtyping relationship with either \C or
\D. Moreover, ifcgen makes no attempt to produce correct method bodies.

\begin{figure}[h]
\begin{lstlisting}
class F { m(x:D):D { x } m(x:*):* { <*>this.m(<D>x) }
          n(x:C):C { x } n(x:*):* { <*>this.n(<C>x) }}
\end{lstlisting}
\caption{wrap result}
\label{fig:monwrap}
\end{figure}


To fix this, we need to add additional guard methods that ensure the
presence of an untyped target method even if cast. The meet of classes \C
and \D in our system is therefore class \xt{F} as shown in
figure~\ref{fig:monwrap}, which has both the typed and untyped methods
specified in both \C and \D. Class \xt{F} is a proper structural subtype of
both \C and \D, and its methods perform semantically correct operations. It
is generated by the wrap metafunction, which produces an equivalent class
from source and target classes. It ensures that any untyped method in the
source is retained in the target, and that typed methods are inserted.

\begin{figure*}[!t]
\center

\begin{center}
\begin{minipage}{10cm}
\begin{align*}
\TRG[M]{\x}\Env &= \x & \\
\TRG[M]{\Call{\e_1}\m{{\e_2}}}\Env &= \DynCall{\ep_1}\m{\ep_2} & 
   \IF\EM{\TypeCk{\K,\Env}{\e[1]}\any ~ \AND ~
    \TRG[M]{\e_1}\Env = \ep_1}
   \EM{~\AND~ \TAG[M]{\e_2}\Env{\any} = \ep_2}\\
\TRG[M]{\Call{\e_1}\m{{\e_2}}}\Env &= \KCall{\ep_1}\m{\ep_2}{\D[1]}{\D[2]} & 
    \IF\EM{\TypeCk{\K,\Env}{\e[1]}\C ~ \AND\; ~ 
    \Mtype\m{\D[1]}{\D[2]}\In\App\K\C}
  ~\EM{\AND ~ \TRG[M]{\e_1}\Env= \ep_1 
   ~\AND~ \TAG[M]{\e_2}\Env{\D[1]} = \ep_2}\\
\TRG[M]{\New\C{{\e_1}..}}\Env &= \New\C{\ep_1..} & 
   \IF\EM{\Ftype{\f[1]}{\t[1]}\In\App\K\C ~\AND~
     \ep[1] = \TAG[M]{\e[1]}\Env{\t[1]} ~..}\\
\TAG[M]{\e}{\Env}\t &= \e' & \IF\EM{\TypeCk{\K,\Env}\e\tp 
     ~\AND~ \K\vdash\tp\Sub\t}\\
\TAG[M]{\e}{\Env}\t &= \MonCast{\t}{\ep} & \IF\EM{\TypeCk{\K,\Env}\e\tp 
    ~\AND~ \K\vdash\tp\not\Sub\t}
\end{align*}
\end{minipage}
\end{center}
\caption{Monotonic Translation}
\end{figure*}
\begin{figure*}
\begin{mathpar}
\IRule{CM}{
  \s(\a) = \C\{\ap..\} \\
  \tmeet \C\t\cdot\K = \Cp~\Kp\\
  \sp = \s[\a \mapsto \Cp\{\ap..\}]
}{
  \moncast \a\t\s\K \Kp \sp\\
}
\end{mathpar}

\begin{mathpar}

\IRule{TM1}{ }{\tmeet\C\any\P\K = \C\,\K}

\IRule{TM2}{ }{\tmeet\any\C\P\K = \C\,\K}

\IRule{TM3}{ }{\tmeet\t\t\P\K = \t\,\K}

\IRule{TM5}{
    \P(\C,\D) = \Cp
}{
    \tmeet\C\D\P\K = \Cp\,\K
}

\IRule{TM4}{
  \fresh\Cp\\
  (\C,\D) \not\in\P \\
  \Pp = \P~{{(\C,\D)}\mapsto\Cp} \\
  \mmeet{\xt{mtypes}(\C,\K)}{\xt{mtypes}(\D,\K)}\Pp\K = \mtpp..\,\Kp\\
  \Kpp = \Kp~\typegen{\mtpp..}\Cp\\
}{
    \tmeet\C\D\P\K = \Cp\,\Kpp
}

\end{mathpar}


\begin{mathpar}
\IRule{MM1}{}{
  \mmeet{\mt ..}{\cdot}\Env\K =\mt.. ~\K
}

\IRule{MM2}{}{
  \mmeet{\cdot}{\mt ..}\Env\K ={\mt ..} ~\K
}

\IRule{MM6}{
  \Mtype\m{\t_3}{\t_4} \in {\mt_2} \\
  \tmeet{\t_3}{\t_1}\Env\K = \t_5~\Kp \\
  \tmeet{\t_2}{\t_4}\Env\Kp = {\t_6}~{\Kpp} \\
  \mmeet{{\mt_1 ..}}{{\mt_2 ..}}\Env\Kp = {\mt_3 ..}~\Kpp
}{
  \mmeet{\Mtype\m{\t_1}{\t_2}~{\mt_1..}}{{\mt_2..}}\Env\K =\Mtype\m %%%TODO: this used to be \n CHECK THAT THIS IS CORRECT
   {\t_5}{\t_6} ~{\mt_3 ..} ~\Kpp
}
\end{mathpar}

\vspace{5mm}

\begin{center}\begin{minipage}{12cm}
\noindent\[\begin{array}{l@{~}r@{}r@{~}r@{~}r@{~}rr}
\arrayrulecolor{white}
\typegen{\mt..}{\D} = \\
\SP \class~\D~\{
\\[1mm]
\SPP \Mdef\m\x\t\tp {{\MonCast\tp{\x}}} 
&
\qquad \All m \Mtype\m\t\tp\in\mt..
\\[1mm]
\SP\}
\end{array}
\]

\vspace{5mm}

\[\begin{array}{l@{~}r@{}r@{~}r@{~}r@{~}rr}
\arrayrulecolor{white}
\classgen{\C, \md.., \mt.., \mtp.., \D, \K}= \\
  \SP \class~\D~\{ \\
    \SPP \Fdef\f\t   &
        \All{\f\!:\t\in\text{fields}(\K,\C)} \\[1mm]
    \SPP \Mdef\m\x\any\any{~\MonCast\any{\KCall\this\m{\MonCast{\t[1]}\x}{\t[1]}{\t[2]}}~}\HS\HS & 
          \All\m \Mdef\m\x{\any}{\any}\e\in\md..  ~\AND~ \Mtype\m{\C_1}{\C_2}\in{\mtp..}
\\[1mm]\hline
\SPP \Mdef\m\x{\C_1}{\C_2} {\MonCast{\C_2}{[{(\MonCast{\any}\x)}/\x]\e}~}
&     \All \m \Mdef\m\x{\any}{\any}\e\in\md.. ~\AND~ \Mtype\m{\C_1}{\C_2}\in\mtp.. 
\\[1mm]\hline
\SPP \Mdef\m\x{\C_1}{\C_2} {\e}~
&     \All \m \Mdef\m\x{\C_1}{\C_2}\e\in\md.. ~\AND~ \Mtype\m{\C_1}{\C_2}\in\mtp.. 
\\[1mm]\hline
\SPP \Mdef\m\x{\any}{\any} {\e}~
&     \All \m \Mdef\m\x{\any}{\any}\e\in\md.. ~\AND~ \Mtype\m{\C_1}{\C_2}\not\in\mtp.. 
\\[1mm]\hline
\SP\}
\end{array}
\]\end{minipage}\end{center}
\caption{Monotonic Cast Operations}
\label{moncast}
\end{figure*}

\section{Formal Example}

To illustrate the operation of our semantics, we present a fully worked
example of a simple program running under the monotonic semantics. This
program, shown in figure~\ref{fig:monoex}, is first translated from source to
\kafka and then executed according to the \kafka semantics described above.

This program is designed to illustrate the cast semantics of monotonic. The
method \xt{alias} aliases its argument, casting it to both \C and \D. The
value cast to \C is then retrieved and its method \xt{check} is then called
(in untyped code) with an \xt{I}. The reference is identical to the one
originally passed to \xt{alias} and no wrapper has been applied. One would
expect then that the untyped call to \xt{check} would pass, since the body of
\xt{check} is the identity function, but under the monotonic semantics it
fails.

Before explaining the semantics of this program in more detail, we will first
describe the top level type-driven transformation. First, casts are inserted
wherever a statically typed value is passed to any and vice versa. Then, typed
methods have untyped shims generated, so they can be called in a protected
manner from untyped code. 

Execution starts with the expression \code{new E().alias(<*>new
C()).check(<*>new C())}. \code{new E()} evaluates to a reference \a to an
object $\xt{E}\{\}$, upon which \xt{alias} is invoked with argument \New\C{}
(which evaluates to \a[1], mapping to $\C\{\}$). \xt{alias} then expands to
\MonCast\any{\New{\xt{T}}{\MonCast\C\a[1],\MonCast\D\a[1]}}. The first
\MonCast\C\a[1] is a no-op, as \a[1] is already of type \C, but the second
\MonCast\D\a[1] modifies the runtime type of the value stored at \a[1].

\D defines \xt{check} to take and return instances of \xt{I}, while \C  has
\xt{check} taking and returning values of type \any. As \D's definition is
more precise, it will take priority. As a result, the result of meet will be
\Cp, with \Class\Cp{}{\Mdef{\xt{check}}\x{\xt{I}}{\xt{I}}\x} appended to the
class table. monWrap is then applied, producing the final new class \Cpp from
\Cp, consisting of \Class\Cpp{}{\Mdef{\xt{check}}\x{\xt{I}}{\xt{I}}{\MonCast{\xt{I}}{\MonCast{\any}{\x}}} ~ \Mdef{\xt{check}}\x\any\any{\MonCast{\any}{\KCall\this{\xt{check}}{\MonCast{\xt{I}}{\x}}{\xt{I}}{\xt{I}} }}}. This is then used to overrite \a[1]'s class in the heap. \a[1] originally
referred to $\C\{\}$, but it will now refer to $\Cpp\{\}$ after the cast.

At this point, the initial expression has been reduced to
\DynCall{\MonCast\any{\KCall{\New{\xt{T}}{\a[1], \a[1]}}{\xt{first}}\any\any
}}{\xt{check}}{\MonCast\any{\New\C{}}}. The inner expression
\KCall{\New{\xt{T}}{\a[1], \a[1]}}{\xt{first}}{\a[1]}\any\any evaluates to
\KCall{\a[2]}{\xt{first}}{\a[1]}\any\any where \a[2] refers to $\xt{T}\{\a[1],\a[1]\}$
in the heap. The \xt{first} function then is called, returning \a[1] under type
\any.

This leaves us with \DynCall{\a[1]}{\xt{check}}{\MonCast\any{\New\C{}}}. We
evaluate \New\C{} to \a[3] (mapped to $\C\{\}$), and make the invocation. A
method \xt{check} under type \any and \any exists, so the program does not get
stuck. \app is an instance of \Cpp, so it is dispatched to the generated
method \xt{check} made by the monocast operation. As a result, our new term is
\MonCast{\any}{\KCall{\a[1]}{\xt{check}}{\MonCast{\xt{I}}{\a[3]}}{\xt{I}}{\xt{I}}}.
However, at this point, we get stuck, as no meet exists between the runtime type of
\a[3] (mapped to $\C\{\}$) and \xt{I}. 
 
\begin{figure}[!h]
\begin{tabular}{cc}
\begin{lstlisting}
class C { check(x:*) { x } }
class D { check(x:I):I { x } }



class I { dif(x:*):* { x }}

class T { 
  f1:C f2:D 
  first(x:*):* { this.f1 } }
class E { 
  alias(x:*):* { 
    new T(x,x).first(x) }}

new E().alias(new C())
       .check(new C())
\end{lstlisting}  &
\begin{lstlisting}
class C { check(x:*) { x } }
class D { check(x:I):I { x }
  check(x:*):* { 
    <*>this.m(<I>x) } }

class I { dif(x:*):* { x }}

class T { 
  f1:C f2:D 
  first(x:*):* { <*>this.f1 } }
class E { 
  alias(x:*):* { 
    <*>new T(<C>x,<D>x).first(x) }}

new E().alias(<*>new C())
       @check(<*>new C())
\end{lstlisting}\\
Source & \kafka 
\end{tabular}
\caption{Monotonic Semantics Example}
\label{fig:monoex}
\end{figure}

\footnotesize
\normalsize

\bibliographystyle{ACM-Reference-Format}
\bibliography{../../bib/jv,../../bib/all}

\end{document}
